ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-02 rebuttal.tex

\begin{response}
    %
    We agree with the reviewer that some of the choices currently enforced in the PyKED package and
    the ChemKED schema do not lend themselves to easy informal use. We presently recommend that
    users who wish to use ChemKED files in this way include an incomplete reference field. We have
    modified the ChemKED schema in response to this comment so that the only required subfields of
    the reference field are the authors and the year.

    In addition, there is a \verb|skip_validation| argument to the \verb|ChemKED| class constructor.
    This option turns off all of the validation, including the validation of the reference material
    and enforcement of required fields. We recognize that this is a limitation at present, if users
    wish to, for example, validate that their units are correct but do not wish to validate some
    other aspect of the file. Work is ongoing to improve this situation, and discussion is contained
    in a public issue on GitHub: \url{https://github.com/pr-omethe-us/PyKED/issues/39}
    %
\end{response}

\begin{reviewer}
    %
    10.~Page 4, line 32: ``The second section of the file encodes the experimental data''. Perhaps
    it is easier to follow if this is briefly mentioned before the authors explain the details of
    the first section of the ChemKED format.
    %
\end{reviewer}

\begin{response}
    %
    We have updated the text to incorporate the suggested change.
    %
\end{response}

@@

    Users can create Excel or CSV files containing all or some of the datapoints in the ChemKED
    instance by using the \verb|get_dataframe()| function, and the associated functionality in
    Pandas (\verb|DataFrame.to_csv()| and \verb|DataFrame.to_excel()|). Since this functionality
    already exists, we do not wish to reinvent the wheel by writing our own functions that would
    have to be maintained and tested. Finally, we would note that data can also be plotted and
    viewed directly from Python using existing functionality in the Pandas DataFrame class
    (\verb|DataFrame.plot()|).
    %
\end{response}

\subsection*{Usage examples}
\subsubsection*{RCM modelling with varying reactor volume}

\begin{reviewer}
    %
    23.~Page 10, line 29--30: ``A user might want to load the information'' Avoid the term `might'.
    It's one of the main goals of using ChemKED and its associated Python-package PyKED to load the
    information necessary for simulating the experiment.
    %
\end{reviewer}

\begin{response}
    %
    We have modified the text to describe this example as an application.
    %
\end{response}

@@

\begin{response}

\end{response}

\begin{reviewer}
    %
    26.~Page 11, line 7: Where is this `air.xml' file coming from. This is not clear from the text.
    %
\end{reviewer}

\begin{response}
    %
    \verb|air.xml| is an input file included with Cantera, representing an average composition of
    air. We have modified the text to include this information.
    %
\end{response}

### 2017-10-02 pyked.tex
Respond to #27

The \yabox|value| can be a string, integer, or floating-point number. We designed
the \ck{} format to include all information necessary to simulate a
given experiment. A \ck{} file is generally broken into two main sections.
The first section details all the ``meta'' information about the experiment
and the \ck{} file itself. This section uses the common \yabox|author| mapping
for any fields that specify an author. The \yabox|author| mapping
contains the following fields:
%
\begin{itemize}
    \item \yabox|name| (required, string): The author's full name
    \item \yabox|ORCID| (optional, string): the author's unique ORCID code
\end{itemize}
%
More generally for the meta information, the required keys in this section
are:
%
\begin{itemize}
    \item \yabox|file-author| (required, \yabox|author|-type mapping): The
    author of the \ck{} file
    \item \yabox|file-version| (required, integer): The version of the \ck{}
    file
    \item \yabox|chemked-version| (required, string): The version of the \ck{}
    schema to which this file conforms
    \item \yabox|reference| (required, mapping): The reference information for
    the published article associated with the data in the file. The fields in this
    mapping are:
    \begin{itemize}
        \item \yabox|authors| (required, sequence): A sequence of \yabox|author|
        mappings
        \item \yabox|journal| (optional, string): The name of the publishing
        journal
        \item \yabox|year| (required, integer): The year of publication
        \item \yabox|doi| (optional, string): The article DOI
        \item \yabox|volume| (optional, integer): The journal volume number
        \item \yabox|pages| (optional, string): The article pages
        \item \yabox|detail| (optional, string): A description of from where the
        data originated (e.g., figure or table number)
    \end{itemize}
    \item \yabox|experiment-type| (required, string): Currently, only
    \yabox|ignition delay| is supported
    \item \yabox|apparatus| (required, string): The type of apparatus used to
    perform the experiment (\yabox|shock tube| or \yabox|rapid compression machine|)
\end{itemize}

The second section of the file encodes the experimental data. The \ck{} file
describes these using a sequence of mappings, with the following information required
in each element of the sequence:
%
\begin{itemize}
    \item \yabox|temperature| (required, sequence): The temperature of the
    experiment, with units and optionally uncertainty
    \item \yabox|ignition-delay| (required, sequence): The ignition delay of the
    experiment, with units and optionally uncertainty
    \item \yabox|pressure| (required, sequence): The pressure of the experiment,
    with units and optionally uncertainty
    \item \yabox|composition| (required, mapping): The composition of the
    mixture in the experiment, described via a nested mapping with the fields
    \begin{itemize}
        \item \yabox|kind| (required, string): one of \yabox|mole fraction|, \yabox|mole percent|, or \yabox|mass fraction|
        \item \yabox|species|: sequence of mappings, with the fields
        \begin{itemize}
            \item \yabox|species-name| (required, string): The name of the species
            \item \yabox|amount| (required, sequence): The mole fraction or percent, or the mass fraction, optionally with uncertainty
            \item \yabox|InChI|, \yabox|SMILES|, \yabox|atomic-composition|, or
            \yabox|elemental-composition| (required, string or sequence): The InChI
            or SMILES string representing the molecule, or its atomic or elemental
            composition (\yabox|atomic-composition| and\\
            \yabox|elemental-composition| are synonyms) as a sequence
        \end{itemize}
    \end{itemize}
    \item \yabox|ignition-type| (required, mapping): The method used to detect
    ignition in the experiments. The required fields are
    \begin{itemize}
        \item \yabox|type| (required, string): How ignition delay was
        measured; one of \yabox|d/dt max| (indicates the ignition point was
        found at the maximum of the time derivative of the \yabox|target|),
        \yabox|max| or \yabox|min| (indicates the ignition point was at the
        maximum or minimum of the \yabox|target|), \yabox|1/2 max| (
        indicates the half-maximum point of \yabox|target|), or \yabox|d/dt max extrapolated|
        (indicates the maximum slope of the target extrapolated to the baseline)
        \item \yabox|target| (required, string): The target for the ignition
        \yabox|type| measurement; one of \yabox|temperature|, \yabox|pressure|,
        \yabox|OH|, \yabox|OH*|, \yabox|CH|, or \yabox|CH*|
    \end{itemize}
\end{itemize}

>>>>

The \yabox|value| can be a string, integer, or floating-point number. We designed
the \ck{} format to include all information necessary to simulate a
given experiment. A \ck{} file is generally broken into two main sections: a section containing the
``meta'' information about the experiment and the \ck{} file itself, and a section containing the
actual experimental data to be encoded.

\subsection{Apparatus Information and Metadata}\label{sec:apparatus-information-and-metadata}

We will first describe the fields in the ``meta'' section.
This section uses a mapping that is called the \yabox|author| mapping in the schema. As this mapping
is used for any fields that specify an author throughout the schema, we describe it first.
The \yabox|author| mapping contains the following fields:
%
\begin{itemize}
    \item \yabox|name| (required, string): The author's full name
    \item \yabox|ORCID| (optional, string): the author's unique ORCID code
\end{itemize}
%
In general, the meta information describes the experimental facility and type of experiment, the
reference from which the data was taken, and the author of the \ck{} file itself. For each of the
keys below, we have indicated whether the key is required or optional, and the type of value
associated with the key. The keys in the meta section include:
%
\begin{itemize}
    \item \yabox|apparatus| (required, mapping): Information about the specific experimental
    apparatus used to conduct the experiment. The fields in this mapping are:
    \begin{itemize}
        \item \yabox|facility| (optional, string): A unique name or identifier for the apparatus,
        if the institution has several that are similar
        \item \yabox|institution| (optional, string): The institution where the experimental
        apparatus was located when the experiments were performed
        \item \yabox|kind| (required, string): The type of apparatus used to
        perform the experiment. Currently, only \yabox|shock tube| or
        \yabox|rapid compression machine| are supported.
    \end{itemize}
    \item \yabox|chemked-version| (required, string): The version of the \ck{}
    schema to which this file conforms
    \item \yabox|experiment-type| (required, string): Currently, only
    \yabox|ignition delay| is supported
    \item \yabox|file-author| (required, \yabox|author|-type mapping): The
    author of the \ck{} file
    \item \yabox|file-version| (required, integer): The version of the \ck{}
    file
    \item \yabox|reference| (required, mapping): The reference information for
    the published article associated with the data in the file. The fields in this
    mapping are:
    \begin{itemize}
        \item \yabox|authors| (required, sequence): A sequence of \yabox|author|
        mappings
        \item \yabox|detail| (optional, string): A description of from where the
        data originated (e.g., figure or table number)
        \item \yabox|doi| (optional, string): The article DOI
        \item \yabox|journal| (optional, string): The name of the publishing
        journal
        \item \yabox|pages| (optional, string): The article pages
        \item \yabox|volume| (optional, integer): The journal volume number
        \item \yabox|year| (required, integer): The year of publication
    \end{itemize}
\end{itemize}

Several of these keys bear further discussion. The \yabox|chemked-version| is present in the file so
that PyKED can determine whether the file conforms to a supported version of the schema. Although
quite similar in name, the \yabox|file-version| represents the number of versions that have been
created of the current file; it is an integer that should be incremented whenever changes are made
to the file. Finally, the \yabox|reference| mapping is required in every file, even those that
represent data that has not been published. However, the \yabox|reference| mapping only requires the
\yabox|authors| sequence and the \yabox|year|, so that \ck{} files can be used internally within
research groups to represent unpublished data prior to publication.

\subsection{Ignition Delay Experimental Data}\label{sec:ignition-delay-experimental-data}

The second section of the file encodes the experimental data as a sequence of mappings. The
top-level key for the sequence is called \yabox|datapoints|. For the current version of the \ck{}
schema, only autoignition delay experiments are supported. The following information os required in
each element of the sequence:
%
\begin{itemize}
    \item \yabox|temperature| (required, sequence): The temperature of the
    experiment, with units and optionally uncertainty
    \item \yabox|ignition-delay| (required, sequence): The ignition delay of the
    experiment, with units and optionally uncertainty
    \item \yabox|pressure| (required, sequence): The pressure of the experiment,
    with units and optionally uncertainty
    \item \yabox|composition| (required, mapping): The composition of the
    mixture in the experiment, described via a nested mapping with the fields
    \begin{itemize}
        \item \yabox|kind| (required, string): one of \yabox|mole fraction|, \yabox|mole percent|, or \yabox|mass fraction|
        \item \yabox|species|: sequence of mappings, with the fields
        \begin{itemize}
            \item \yabox|species-name| (required, string): The name of the species
            \item \yabox|amount| (required, sequence): The mole fraction or percent, or the mass fraction, optionally with uncertainty
            \item \yabox|InChI|, \yabox|SMILES|, \yabox|atomic-composition|, or
            \yabox|elemental-composition| (required, string or sequence): The InChI
            or SMILES string representing the molecule, or its atomic or elemental
            composition (\yabox|atomic-composition| and\\
            \yabox|elemental-composition| are synonyms) as a sequence
        \end{itemize}
    \end{itemize}
    \item \yabox|ignition-type| (required, mapping): The method used to detect
    ignition in the experiments. The required fields are
    \begin{itemize}
        \item \yabox|type| (required, string): How ignition delay was
        measured; one of \yabox|d/dt max| (indicates the ignition point was
        found at the maximum of the time derivative of the \yabox|target|),
        \yabox|max| or \yabox|min| (indicates the ignition point was at the
        maximum or minimum of the \yabox|target|), \yabox|1/2 max| (
        indicates the half-maximum point of \yabox|target|), or \yabox|d/dt max extrapolated|
        (indicates the maximum slope of the target extrapolated to the baseline)
        \item \yabox|target| (required, string): The target for the ignition
        \yabox|type| measurement; one of \yabox|temperature|, \yabox|pressure|,
        \yabox|OH|, \yabox|OH*|, \yabox|CH|, or \yabox|CH*|
    \end{itemize}
\end{itemize}
